### 2017-09-25 aaV_main.tex
added main file

\documentclass[12pt]{article}

%\textwidth=20cm
%\textheight=16cm
%\setlength{\oddsidemargin}{-1cm}
\pdfoutput=1
\setlength\parindent{0pt}

\usepackage{jheppub}
\usepackage{graphicx,epsfig,amsmath,amssymb,booktabs}
%\usepackage{graphicx,epsfig,amsmath,amssymb}
%\usepackage{multirow,bigdelim,lscape}
%\usepackage{caption}
\usepackage{caption,subcaption}
\usepackage{xspace}

%\usepackage{listings}
%\lstset{basicstyle=\ttfamily}
%% 
%\usepackage{titlesec}

%%% Macros
\newcommand{\eqn}[1]{Eq.~(\ref{#1})}
\newcommand{\ab}{\ensuremath{\alpha_0}}
\newcommand{\als}{\ensuremath{\alpha_s}}
\newcommand{\aapa}{\ensuremath{\gamma\gamma}}
\newcommand{\shortequal}{\ensuremath{\!\!\!=\!\!\!}}
\newcommand{\pT}{\ensuremath{p_\text{T}}}
\newcommand{\order}{\ensuremath{\mathcal{O}}}

\newcommand{\Sherpa}{S\scalebox{0.8}{HERPA}\xspace}
\newcommand{\GoSam}{G\scalebox{0.8}{O}S\scalebox{0.8}{AM}\xspace}
\newcommand{\Recola}{R\scalebox{0.8}{ECOLA}\xspace}
\newcommand{\OpenLoops}{O\scalebox{0.8}{PEN}L\scalebox{0.8}{OOPS	}\xspace}
\newcommand{\QGraf}{QG\scalebox{0.8}{RAF}\xspace}
\newcommand{\FORM}{F\scalebox{0.8}{ORM}\xspace}
\newcommand{\Spinney}{S\scalebox{0.8}{PINNEY}\xspace}
\newcommand{\Ninja}{N\scalebox{0.8}{INJA}\xspace}
\newcommand{\Samurai}{S\scalebox{0.8}{AMURAI}\xspace}
\newcommand{\GolemNF}{G\scalebox{0.8}{OLEM}95\xspace}
\newcommand{\OneLoop}{O\scalebox{0.8}{NE}L\scalebox{0.8}{OOP}\xspace}
\newcommand{\FastJet}{F\scalebox{0.8}{AST}J\scalebox{0.8}{ET}\xspace}

\newcommand{\DIPHOX}{D\scalebox{0.8}{IPHOX}\xspace}
\newcommand{\MCFM}{M\scalebox{0.8}{CFM}\xspace}
\newcommand{\twogammaMC}{2\scalebox{0.8}{gammaMC}\xspace}
\newcommand{\VBFatNLO}{V\scalebox{0.8}{BF@NLO}\xspace}
\newcommand{\RESBOS}{R\scalebox{0.8}{ESBOS}\xspace}

\newcommand{\comment}[1]{\textbf{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{frontmatter}

\title{Electroweak corrections to diphoton plus Vector Boson}

\author[a]{Nicolas Greiner,}
\author[a,b]{Marek Sch\"onherr}

\affiliation[a]{Physik Institut, Universit{\"a}t Z{\"u}rich, Winterthurerstr.190, 8057 Z\"urich, Switzerland}
\affiliation[b]{CERN}

\preprint{
  \small
  \begin{flushright}
    ZU--TH 
  \end{flushright}
}

\abstract{
  We calculate the next-to-leading order electroweak corrections...
}

\keywords{EW corrections, Photon, NLO, Jets}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Introduction
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Calculational Setup
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Calculational setup}
\label{sec:setup}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Results
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\input{results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Conclusions
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}

\section*{Acknowledgements}
N.G.\ was supported by the Swiss National Science Foundation under contract
PZ00P2\_154829. M.S.\ was supported by PITN--GA--2012--315877 ({\it MCnet}) 
and the ERC Advanced Grant MC@NNLO (340983).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\providecommand{\href}[2]{#2}\begingroup\raggedright\begin{thebibliography}{100}

\end{thebibliography}\endgroup

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%