### 2017-10-03 conclusions.tex
restructured

\section{Conclusions}
\label{sec:conclusions}

We did good.