### 2017-10-03 intro.tex

\section{Introduction}
\label{sec:intro}

We love to calculate things that are important.
\cite{Gleisberg:2008ta}