### 2017-10-03 setup.tex

\section{Calculational setup}
\label{sec:setup}

We do it in a way that works.