### 2017-10-17 intro.tex
update intro and merge

\section{Introduction}
\label{sec:intro}

The precise understanding of the electroweak symmetry breaking mechanism is an important cornerstone in verifying the
Standard Model. Different realizations of the electroweak sector or new physics coupling to electroweak gauge bosons will
yield to deviations compared to the Standard Model prediction. New physics effects can conveniently be described in terms
of an effective theory where new heavy degrees of freedom are integrated out and deviations from the Standard Model are 
parametrised by higher dimensional operators \cite{Weinberg:1978kz,Weinberg:1979pi}. Higher dimensional operators can lead
 to deviations in triple and quartic gauge couplings. Moreover, new vertices (e.g. $Z\gamma\gamma$) that do not exist in the 
Standard Model can appear. The class of processes that involve a pair of photons in association with another vector boson
allows to measure deviations in triple and quartic gauge couplings and is therefore a particularly interesting class of processes.
 Consequently both ATLAS and CMS have measured these type of processes und derived constraints on anomalous gauge
 couplings \cite{Aad:2016sau,Sirunyan:2017lvq,Aad:2015uqa,Aad:2015bua}