### 2017-12-05 RefereeReply1.txt
corrections for points 8, 9, 10, 11 and 13 of referee report, added draft reply

We would like to thank the referee for his/her constructive and valuable 
comments.  The points raised clearly helped us to further improve the 
quality of the manuscript.  Below we give detailed replies to the referee's 
queries and indicate the corresponding adjustments or additions to the 
manuscript.  We thereby follow the referee's initial ordering: 
1) 
2)
3)
4)
5)
6)
7)
8)  The figure captions have been corrected.
9)  This has been rephrased to clearly state which particles the invariant 
    mass cut acts on.
10) The PDFs have been specified, and the respective publication referenced.
11) 
12) The description of the observables has been enhanced in the following 
    cases:
    - 
    - 
13) We agree and added a sentence stating exactly that.

We again would like to thank for the his/her feedback that clearly helped us to 
further improve the quality of the paper. We hope to have addressed all raised 
questions and suggestions such that the paper in its present form can be 
published in JHEP.

Sincerely yours

Nicolas Greiner, Marek Schoenherr