### 2017-12-05 RefereeReply1.txt
corrected bib, added reply to 6) and 7)

We would like to thank the referee for his/her constructive and valuable 
comments.  The points raised clearly helped us to further improve the 
quality of the manuscript.  Below we give detailed replies to the referee's 
queries and indicate the corresponding adjustments or additions to the 
manuscript.  We thereby follow the referee's initial ordering: 
1) 
2)
3)
4)
5)  We have somewhat improved the plots to better distinguish the 
    results with and without a jet veto. The caption of Fig. 4 is 
    amended accordingly. The focus, however, is on the ratios where 
    only a subset of results is plotted and legibility is, in our 
    estimation, sufficient for the points raised in the manuscript.
6)  As detailed in the manuscript in the paragraph above eq. (3.3), 
    we compute the LO result including its factorisation scale 
    uncertainty, and the EW correction factor for the central scale 
    choice. The NLO EW result, according to eq. (3.3) thus carries 
    the LO factorisation scale uncertainty.
    As labeled, the plots themselves show not delta-EW, which we compute only 
    for the central scale choice, but the full NLO EW result exhibiting 
    the LO factorisation scale uncertainty. In the ratios the NLO EW 
    result in relation to the central LO prediction is shown, i.e. 
    the resulting uncertainty band has the same relative size as 
    the LO uncertainty band.
    We have amended the caption of Fig. 4 to explicitly reflect 
    the reference of the ratio plots.
7)  We have rendered this statement more precise and hope that it 
    is clear now.
8)  The figure captions have been corrected.
9)  This has been rephrased to clearly state which particles the invariant 
    mass cut acts on.
10) The PDFs have been specified, and the respective publication referenced.
11) 
12) The description of the observables has been enhanced in the following 
    cases:
    - 
    - 
13) We agree and added a sentence stating exactly that.

We again would like to thank for the his/her feedback that clearly helped us to 
further improve the quality of the paper. We hope to have addressed all raised 
questions and suggestions such that the paper in its present form can be 
published in JHEP.

Sincerely yours

Nicolas Greiner, Marek Schoenherr